\textit{Supplemental materials for}: How is experience transformed into memory?

### 2018-09-04 paper/supplemental_materials.tex
backing up progress on supp mats [WIP]

\paragraph{Temporal clustering scores.} Temporal clustering refers to the extent to which participants group their recall responses according to encoding position~\citep{PolyEtal09}. For instance, if the participant recalled the video events in the exact order they occurred, this would yield a score of 1.  If the participant recalled the video events in reverse order, this would yield a score of 0.  Similarly, if the participant recalled the events in random order, this yield yield an expected score of 0.5.  For each recall event transition (and separately for each participant), we sorted all not-yet-recalled events according to their absolute lag (i.e., distance away in the video).  We then computed the percentile rank of the next event the participant recalled.  We averaged these percentile ranks across all of the participant's recalls to obtain a single temporal clustering score for the participant.  Overall, we found that participants with higher temporal clustering scores also tended to recall more events ($r = 0.62, p = 0.007$).

\paragraph{Semantic clustering scores.} Semantic clustering measures the extent to which participants grouped their recall responses according to semantic similarity~\citep{PolyEtal09}. Here, we used the topic vectors for each event as a proxy for its semantic content. Thus, the similarity between the semantic content for two events can be computed by correlating their respective topic vectors.  For each recall event transition, we sorted all not-yet-recalled events according to how correlated the topic vector \textit{of the closest-matching video event} was to the topic vector of the closest-matching video event to the just-recalled event.  We then computed the percentile rank of the observed next recall.  We averaged these percentile ranks across all of the participant's recalls to obtain a single semantic clustering score for the participant.  We found that participants who exhibited stronger semantic clustering scores overall remembered more video events ($r = 0.55, p = 0.02$).

@@

\subsection*{Additional measures of naturalistic memory}
To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the video model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered and computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events out of order, or with low precision, the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p < 0.05$, participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to hand annotated memory performance across participants (Pearson's $r(16) =-.57, p = 0.016$). Thus, a closer match between the video and recall event models was related to better recall performance.

Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching video event (i.e the ``precision'' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between hand annotated memory performance and precision suggesting that participants who remembered more events also remembered them more veridically (Pearson's $r(16) = 0.74, p = 0.0006$).

Then, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e., 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher hand annotated memory performance also had higher distinctiveness scores (Pearson's $r(16) = 0.8, p = 0.0001$).

>>>>

\paragraph{Temporal clustering scores.} Temporal clustering refers to the extent to which participants group their recall responses according to encoding position~\citep{PolyEtal09}. For instance, if the participant recalled the video events in the exact order they occurred, this would yield a score of 1.  If the participant recalled the video events in reverse order, this would yield a score of 0.  Similarly, if the participant recalled the events in random order, this yield yield an expected score of 0.5.  For each recall event transition (and separately for each participant), we sorted all not-yet-recalled events according to their absolute lag (i.e., distance away in the video).  We then computed the percentile rank of the next event the participant recalled.  We averaged these percentile ranks across all of the participant's recalls to obtain a single temporal clustering score for the participant (mean: 0.808, SEM: 0.022).  Overall, we found that participants with higher temporal clustering scores also tended to recall more events ($r = 0.62, p = 0.007$).

\paragraph{Semantic clustering scores.} Semantic clustering measures the extent to which participants grouped their recall responses according to semantic similarity~\citep{PolyEtal09}. Here, we used the topic vectors for each event as a proxy for its semantic content. Thus, the similarity between the semantic content for two events can be computed by correlating their respective topic vectors.  For each recall event transition, we sorted all not-yet-recalled events according to how correlated the topic vector \textit{of the closest-matching video event} was to the topic vector of the closest-matching video event to the just-recalled event.  We then computed the percentile rank of the observed next recall.  We averaged these percentile ranks across all of the participant's recalls to obtain a single semantic clustering score for the participant (mean: 0.813, SEM: 0.022).  We found that participants who exhibited stronger semantic clustering scores overall remembered more video events ($r = 0.55, p = 0.02$).

\subsection*{Additional measures of naturalistic memory}
To quantify the similarity between the video topic trajectory and individual recall topic trajectories, we considered several novel metrics.  First, we tested whether each participant's recall trajectory matched the video trajectory in a general sense. For each participant we filtered the video trajectory to only include the events that the participant remembered.  We then computed the root mean squared difference (RMSD) between the remaining video events and the (closest-matching) recalled events.  For example, if the topic vectors for a participant's recall event topic vectors matched the corresponding video event topic vectors exactly (and in order), the expected RMSD for those events would be 0.  However, if the participant's recall events did not perfectly match the video events, or if they were out of order, then the RMSD would be greater than 0.  To assess the significance of the match between the video and recall trajectories, we carried out a permutation procedure whereby, for each of 10000 repetitions, we circularly shifted the recall trajectories (in time) by a random amount and then re-computed the RMSD each time.  This yielded a distribution of ``null'' RMSD values for each participant.  The observed RMSD values reached significance (i.e., $p < 0.05$, reflecting that more than 95\% of the null RMSD values were greater than the observed RMSD value) for nine of the participants (3, 4, 8--13, and 17).  (For the remaining participants this test yielded $p$s$ < 0.25$.)  The observed RMSD values were also reliably correlated with hand-annotated memory performance across participants ($r = -0.57, p = 0.016$).  In other words, a closer match between the video and recall topic trajectories was related to better overall recall performance.

\paragraph*{Precision.}
We tested whether participants who recalled more events were also more \textit{precise} in their recollections. For each participant, we computed the correlation between the topic vectors for each recall event and that of its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event,  indexing how similar the recall event was to its matching video event.  We defined the precision as the average video-recall correlation across all of the events a participant recalled.  We found a strong correlation between hand-annotated memory performance and precision, suggesting that participants who remembered more events also remembered them more veridically ($r= 0.74, p = 0.0006$).

@@

\paragraph*{Distinctiveness.}
We also considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e., 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher hand annotated memory performance also had higher distinctiveness scores (Pearson's $r(16) = 0.8, p = 0.0001$).
